On exponential sums arising from the classical doubling method

### 2025-03-20 paper.tex
Writing about the different groups

\section{Notation}
Let $\finiteField$ be a finite field with $q$ elements, and let $\fieldCharacter \colon \finiteField \to \multiplicativegroup{\cComplex}$ be a non-trivial character. Fix an algebraic closure $\algebraicClosure{\finiteField}$ of $\finiteField$. For any $k \ge 1$, let $\finiteFieldExtension{k} \slash \finiteField$ be the unique field extension of degree $k$ contained in $\algebraicClosure{\finiteField}$.

@@

	\item (Even special orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. Denote $\dim_{\finiteField} \hermitianSpace = 2n$. In this case, we have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det \hermitianSpace = 1$, and that $\GroupExtension{G}$ consists of all elements $g \in \GroupExtension{\IsometryGroup\left(\hermitianSpace\right)}$ such that $\det g = \similitudeCharacter\left(g\right)^n$ \calvin{not sure I get what this $\Tilde{G}$ is. I thought it would just be $\widetilde{Isom(V)}$-or maybe its identity component. As it reads it sounds like you are supposed to fix sim and then define $\tilde{G}$.} \elad{This is the definition used is Lusztig's paper for example. It is the identity component of $\GroupExtension{\IsometryGroup\left(\hermitianSpace\right)}$. Generally speaking, we have the equality $\det g^2 = \similitudeCharacter\left(g\right)^{2n}$ (you can see this by taking the determinant of both sides). So the identity component is the matrices for which the ``square root'' of this equality holds, i.e. $\det g = \similitudeCharacter\left(g\right)^n$ and not $\det g = -\similitudeCharacter\left(g\right)^n$} There are two cases here.

@@

		\item Split case: in this case, we can write $\hermitianSpace = \xIsotropic \oplus \yIsotropic$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In this case, we have that $G$ is isomorphic to the subgroup $\SO_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$ consisting of elements $g$ satisfying $$g \cdot w_{2n} \cdot \transpose{g} = w_{2n}$$ and $\det g = 1$.
		The group $\GroupExtension{G}$ is isomorphic to the subgroup $\GSO_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$, consisting of elements $g$ such that $$g \cdot w_{2n} \cdot \transpose{g} = \similitudeCharacter\left(g\right) w_{2n},$$
		and such that $\det g = \similitudeCharacter\left(g\right)^n$, where $\similitudeCharacter\left(g\right)$ is a non-zero scalar.
		\item Non-split case: \calvin{do we want to do even characteristic? Not sure what the classification of quadratic forms over even char are} \elad{Let's not do characteristic two} for odd characteristic there is one non-split case when $V = X \oplus \mathbb{E} \oplus Y$ where $\mathbb{E}$ is the unique degree two extension of $\mathbb{F}$ with quadratic form $Nm_{\mathbb{E}:\mathbb{F}}$ \calvin{find someone to cite}

@@

	\item (Symplectic groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = -1$. In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ for some positive integer $n$ and $G$ is isomorphic to subgroup $\Sp_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$ consisting of all elements $g$ satisfying $$g \cdot \begin{pmatrix}

@@

	\end{pmatrix}.$$ The group $\GroupExtension{G}$ is is isomorphic to the subgroup $\GSp_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$ consisting of all elements $g$ such that $$g \cdot \begin{pmatrix}

@@

	\end{pmatrix} \cdot \transpose{g} = \similitudeCharacter\left(g\right) \begin{pmatrix}

@@

	\end{pmatrix},$$
	where $\similitudeCharacter\left(g\right)$ is a non-zero scalar.
	\item (Unitary groups): $\quadraticExtension \ne \finiteField$. Denote $\dim_{\quadraticExtension} \hermitianSpace = n$. In this case, the group $G = \GroupExtension{G}$ is isomorphic to the subgroup $\UnitaryGroup_n\left(\finiteField\right)$ of $\GL_{n}\left(\quadraticExtension\right)$ consisting of all elements $g$ such that $$g \cdot w_n \cdot \involution{\left(\transpose{g}\right)} = w_n.$$

>>>>

\section{Notation}
Let $\finiteField$ be a finite field with $q$ elements, where $2 \nmid q$, and let $\fieldCharacter \colon \finiteField \to \multiplicativegroup{\cComplex}$ be a non-trivial character. Fix an algebraic closure $\algebraicClosure{\finiteField}$ of $\finiteField$. For any $k \ge 1$, let $\finiteFieldExtension{k} \slash \finiteField$ be the unique field extension of degree $k$ contained in $\algebraicClosure{\finiteField}$.

@@

	\item (Special even orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. Denote $\dim_{\finiteField} \hermitianSpace = 2n$. It follows by definition that every element $g \in \IsometryGroup\left(\hermitianSpace\right)$ satisfies $\left(\det g\right)^2 = 1$. Similarly, every element $g \in \widetilde{\IsometryGroup\left(\hermitianSpace\right)}$ satisfies $\left(\det g\right)^2 = \similitudeCharacter\left(g\right)^{2n}$. In this case, we obtain the connected components of the identity by adding a restriction on the determinant: We have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det \hermitianSpace = 1$, and that $\GroupExtension{G}$ consists of all elements $g \in \GroupExtension{\IsometryGroup\left(\hermitianSpace\right)}$ such that $\det g = \similitudeCharacter\left(g\right)^n$. There are two cases here.

@@

		\item Split case: in this case, we can write $\hermitianSpace = \xIsotropic \oplus \yIsotropic$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In this case, we have that $G$ is isomorphic to $\SO^{+}_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\SO}^{+}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid g \cdot w_{2n} \cdot \transpose{g} = w_{2n},\,\, \det g = 1\right\}.$$
		The group $\GroupExtension{G}$ is isomorphic to the subgroup $\GSO^{+}_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$, consisting of elements $g$ such that defined as the projection to the first coordinate of the $\finiteField$-points of $$\algebraicGroup{\GSO}^{+}_{2n} = \left\{\left(g, \lambda\right) \in \algebraicGroup{\GL}_{2n} \times \multiplcativeScheme \mid g \cdot w_{2n} \cdot \transpose{g} = \lambda w_{2n},\,\, \det g = \lambda^n\right\}.$$
		\item Non-split case: In this case, there exists a decomposition $$\hermitianSpace = \xIsotropic \oplus \hermitianSpace' \oplus \yIsotropic,$$
		where $\xIsotropic$ and $\yIsotropic$ are totally isotropic spaces in duality and $\hermitianSpace'$ is an antisotropic space (that is, for any $0 \ne v' \in \hermitianSpace'$, we have $\innerproduct{v'}{v'} \ne 0$) of dimension 2. Let $d$ be an element of $\multiplicativegroup{\finiteField}$ that is not a square. Denote $$B_{n,d} = \begin{pmatrix}
			\IdentityMatrix{n-1}\\
			& & 1 &\\
			& -d & &\\
			& & & \IdentityMatrix{n-1}
		\end{pmatrix} \cdot w_{2n}.$$
		In this case, we have that $G$ is isomorphic to $\SO^{-}_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\SO}^{-}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid g \cdot B_{n,d} \cdot \transpose{g} = B_{n,d},\,\, \det g = 1\right\}.$$
		The group $\GroupExtension{G}$ is isomorphic to the subgroup $\GSO^{-}_{2n}\left(\finiteField\right)$ of $\GL_{2n}\left(\finiteField\right)$, consisting of elements $g$ such that defined as the projection to the first coordinate of the $\finiteField$-points of $$\algebraicGroup{\GSO}^{-}_{2n} = \left\{\left(g, \lambda\right) \in \algebraicGroup{\GL}_{2n} \times \multiplcativeScheme \mid g \cdot B_{n,d} \cdot \transpose{g} = \lambda B_{n,d},\,\, \det g = \lambda^n\right\}.$$

@@

	\item (Symplectic groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = -1$. In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ for some positive integer $n$ and $G$ is isomorphic to the group $\Sp_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\Sp}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid g \cdot \begin{pmatrix}

@@

	\end{pmatrix}\right\}.$$ The group $\GroupExtension{G}$ is isomorphic to the group $\GSp_{2n}\left(\finiteField\right) \subset \GL_{2n}\left(\finiteField\right)$, defined as the projection to the first coordinate of the $\finiteField$-points of $$\algebraicGroup{\GSp}_{2n} = \left\{\left(g, \lambda\right) \in \algebraicGroup{\GL}_{2n} \times \multiplcativeScheme \mid g \cdot \begin{pmatrix}

@@

	\end{pmatrix} \cdot \transpose{g} = \lambda \begin{pmatrix}

@@

	\end{pmatrix}\right\}.$$
	\item (Unitary groups): $\quadraticExtension \ne \finiteField$. Denote $\dim_{\quadraticExtension} \hermitianSpace = n$. In this case, the group $G = \GroupExtension{G}$ is isomorphic to the group $\UnitaryGroup_n\left(\finiteField\right) \subset\GL_{n}\left(\quadraticExtension\right)$ defined as the $\finiteField$-points of $$\algebraicGroup{\UnitaryGroup}_n = \left\{ g \in \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{\GL}_n} \mid g \cdot w_n \cdot \involution{\left(\transpose{g}\right)} = w_n \right\}.$$
